---
title: "EDA"
author: "Vinayak Kannan (vk297)"
date: "24/04/2022"
output: html_document
---

```{r}
# Loading in the libraries
library(tidyverse)
library(janitor)
library(lubridate)
library(skimr)
library(knitr)
library(naniar)
library(GGally)
library(styler)
library(ggtext)
library(data.table)
library(grid)
library(timeDate)
library(ggplot2)
library(plotly)
#library(ggplot)
#library(dpylr)
```


```{r}
df <-  read.csv("nyt-articles-2020.csv")
head(df)
```

```{r}
str(df)
```

### creating a month column
```{r}
df$pub_date <- as.Date(df$pub_date)
```
```{r}
df$month = format(df$pub_date,"%m")
```


```{r}
head(df)
```
### Seeing populairty for : Newsdesk / section / material

#### Seeing Newsdesk

Lets first see the total number of comments for each News Desk

```{r}
p4 <- ggplot(df %>%
                group_by(newsdesk) %>%
                summarize_at("n_comments",sum,na.rm=T),
              aes(x = n_comments, y = newsdesk, fill = newsdesk)) +
  xlab("Total Comments") +
  ylab("News Desk") +
  scale_y_discrete(guide = guide_axis(check.overlap = TRUE))+
  ggtitle("Which News Desk had the most comments") +
  # place the legend at the bottom
  theme(legend.position = "none")+
  geom_bar(stat="identity",na.rm=TRUE)
```

```{r}
p4
```
It looks like 'Op Ed" has the most activity on its comments. 
"Weekend" , "Real Estate" and "New york" seem to be other notable News Desks with respect to the number of comments on the articles.

Let us also see how the number of comments for each news desk changes with time.

```{r}
p1 <- ggplot(df %>%
                group_by(month,newsdesk) %>%
                summarize_at("n_comments",sum,na.rm=T),
              aes(month,n_comments,fill = newsdesk),fig(400,20)) +
  xlab("Month") +
  ylab("Total Comments") +
  ggtitle("Monthly Total Comments for each News Desk") +
  # place the legend at the bottom
  theme(legend.position = "none")+
  geom_bar(stat="identity",na.rm=TRUE)
```

```{r}
ggplotly(p1)
```

The trend stays common through the year. 
Op-Ed seems to have the most comments through all the months of the year.

All News Desks seem to stay consistent through the year with regards to their respective range of number of comments.


#### Seeing Section

Lets now see the total number of comments for each Section

```{r}
p5 <- ggplot(df %>%
                group_by(section) %>%
                summarize_at("n_comments",sum,na.rm=T),
              aes(x = n_comments, y = section, fill = section)) +
  xlab("Total Comments") +
  ylab("Section") +
 scale_y_discrete(guide = guide_axis(n.dodge=2))+
  ggtitle("Which Section had the most comments") +
  # place the legend at the bottom
  theme(legend.position = "none")+
  geom_bar(stat="identity",na.rm=TRUE)
```

```{r}
p5
```



We see that most sections in the dataset have no comments at all.
"Opinion" and "US" seem to have the most comments

Let us see the if the trend stays same with time

```{r}
p2 <- ggplot(df %>%
                group_by(month,section) %>%
                summarize_at("n_comments",sum,na.rm=T),
              aes(month,n_comments,fill = section),fig(400,20)) +
  xlab("Month") +
  ylab("Total Comments") +
  ggtitle("Total Comments by Section") +
  # place the legend at the bottom
  theme(legend.position = "none")+
  geom_bar(stat="identity",na.rm=TRUE)
```

```{r}
ggplotly(p2)
```


#### Seeing Material

Lets now see the total number of comments for each Material

```{r}
p6 <- ggplot(df %>%
                group_by(material) %>%
                summarize_at("n_comments",sum,na.rm=T),
              aes(x = n_comments, y = material, fill = material)) +
  xlab("Total Comments") +
  ylab("Material") +
 #scale_y_discrete(guide = guide_axis(n.dodge=2))+
  ggtitle("Which Material had the most comments") +
  # place the legend at the bottom
  theme(legend.position = "none")+
  geom_bar(stat="identity",na.rm=TRUE)
```

```{r}
p6
```

"News" is the branch woth the highest comments.



```{r}
p3 <- ggplot(df %>%
                group_by(month,material) %>%
                summarize_at("n_comments",sum,na.rm=T),
              aes(month,n_comments,fill = material),fig(400,20)) +
  xlab("Month") +
  ylab("Total Comments") +
  ggtitle("Total Comments by Material") +
  # place the legend at the bottom
  theme(legend.position = "none")+
  geom_bar(stat="identity",na.rm=TRUE)
```

```{r}
ggplotly(p3)
```


### Let us Now work with the keywords column

Lets first convert the column into a list
```{r}
## Removing the first and last bracket
library(stringr)
#df$keywords
df$keywords <-  str_sub(df$keywords,2,−2)
```

```{r}
df$keywords <- strsplit(df$keywords, ",")
```

```{r}
head(df)
```

```{r}
unnest(df)
```

```{r}
impwords <- df %>% group_by(keywords) %>%summarise(number = n())
```

```{r}
head(impwords)
```


```{r}
cols <- c("keywords","count")
colnames(impwords) <- cols
```

```{r}
top_words <- impwords[order(-impwords$count),]
```


Lets look at the top 20 words
```{r}
top_words[1:20,]
```
The most talked about topic in the year 2020 seems to be Covid19
This makes sense.
We see that the top 20 keywords are mostly political terms , and terms related to covid.
This could mean that most articles were speaking about how different governments dealt with the pandemic.