---
title: "EDA"
author: "Vinayak Kannan (vk297)"
date: "24/04/2022"
output: html_document
---

```{r}
# Loading in the libraries
library(tidyverse)
library(janitor)
library(lubridate)
library(skimr)
library(knitr)
library(naniar)
library(GGally)
library(styler)
library(ggtext)
library(data.table)
library(grid)
library(timeDate)
library(ggplot2)
library(plotly)
#library(ggplot)
#library(dpylr)
```


```{r}
df <-  read.csv("nyt-articles-2020.csv")
head(df)
```

```{r}
str(df)
```

### creating a month column
```{r}
df$pub_date <- as.Date(df$pub_date)
```
```{r}
df$month = format(df$pub_date,"%m")
```

### Let us Now work with the keywords column

Lets first convert the column into a list
```{r}
## Removing the first and last bracket
library(stringr)
#df$keywords
df$keywords <-  str_sub(df$keywords,2,−2)
```

```{r}
df$keywords <- strsplit(df$keywords, ",")
```

```{r}
head(df)
```

```{r}
df <- unnest(df)
```

```{r}
#df$keywords <- toString(df$keywords)
head(df)
```


```{r}
impwords <- df %>% group_by(keywords) %>%summarise(number = n())
```

```{r}
head(impwords)
```


```{r}
cols <- c("keywords","count")
colnames(impwords) <- cols
```

```{r}
top_words <- impwords[order(-impwords$count),]
```


Lets look at the top 20 words
```{r}
top_words[1:20,]
```
The most talked about topic in the year 2020 seems to be Covid19
This makes sense.
We see that the top 20 keywords are mostly political terms , and terms related to covid.
This could mean that most articles were speaking about how different governments dealt with the pandemic.

```{r}
# Let us remove the first and last character of the keywords column
top_words$keywords <-  str_sub(top_words$keywords,2,−2)
```

```{r}
head(top_words)
```
```{r}
## Looks like we have to remove the first character again, there might have been a space stored in the string 
top_words$keywords <-  str_sub(top_words$keywords,2)
```

```{r}
head(top_words)
```
### Wordcloud

```{r}
library(ggwordcloud)
```

```{r}
df$keywords <- str_sub(df$keywords,3,-2)
head(df)
```

```{r}
p2 <- ggplot(top_words, aes(label = keywords, size = count)) +
  geom_text_wordcloud() +
  theme_minimal()
```

```{r}
p2
```

